---
title: "Usage of the climex web application"
author: "Philipp Müller"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Usage of the climex web application}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

# About
This vignette describes the basic functions of the package's web application. Its main purposes are
- To allow the user to interactively and persistently alter the default preprocessing and fitting options for the maximum likelihood fit of the generalized extreme value (GEV) distribution 
- To apply it consistently to many different time series.
- To verify the validity of the optimization procedure using a variety of goodness of fit plots and an animation of the actual optimization procedure.


# Prerequisites
First of all the global variable **CLIMEX.PATH** has to be set. Just add the following lines to ~/.Rprofile

```{r prerequisites, cache = TRUE}
CLIMEX.PATH <<- "~/R/climex/"
if ( !dir.exists( CLIMEX.PATH ) )
    dir.create( CLIMEX.PATH, recursive = TRUE )
```

Afterwards the web app is invoked locally by calling the *climex()* wrapper function and browsing shown IP (e.g. *http://127.0.0.1:3838*)
```{r wrapper, eval = FALSE, dependson = "prerequisites"}
require( climex )
climex()
```
This wrapper function will copy several assets into the **CLIMEX.PATH** folder which will also contain the images necessary to create the animation.


# leaflet tab
# general tab
# likelihood tab
# deploying the app on server side
